Towards a standardized evaluation of imputation methodology

### 2022-07-21 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
Copy paste sections

\subsection{Data generation problems}

@@

\subsection{Missingness generation problems}

@@

\begin{table}[tb]
\begin{center}
\caption{Missingness mechanisms. [TODO: add refs \citet{seam13, meal15, dore18} for clarification and alternatives outlined in \citet{moha21, scho21, more18}]}
\label{table:mech}
\begin{tabular}{lll}
\hline
      & Interpretation and consequences of the mechanism \\
\hline  
MCAR  & Missing completely at random: The probability to be missing is the same for all \\
      & cases. In other words, the missing values are missing at random and the observed \\
      & values are observed at random. \\ \noalign{\smallskip}
MAR   & Missing at random: The probability to be missing is the same within groups of \\
      & cases defined by the observed data only. The essence is that observed relations in \\
      & the data inform the missingness, such that MCAR may be assumed within the \\ 
      & observed groups. \\ \noalign{\smallskip}
MNAR  & Missing not at random: The probability to be missing depends on unobserved \\
      & aspects of the data. The cause of the missingness is unknown and cannot be \\ 
      & inferred from the observed data. This is considered non-ignorable missingness \\
      & \citep[see e.g.][]{rubi76}. \\
\hline
\end{tabular}
\end{center}
\end{table}

@@

\subsection{Evaluation problems}

@@

\section{Suggested course of action}

@@

\subsection{Set scope}

\subsection{Obtain truth}

\subsection{Induce missingness}

In addition to varying the missingness mechanisms, each simulated mechanism should be combined with different missingness patterns. Remember that missingness is only ignorable under MAR when the parameter of the data is distinct and a-priori independent from the parameter of the missing data process. Under MAR missingness we assume that we may use the observed data to make inferences about the joint (observed and unobserved) data. The dependency of the procedure on the assumption under which we obtain inference is only influenced by the amount of missingness. If there is no missingness--or if there is no data, for that matter--the inference does not depend on the assumption. Alternatively, the validity of assumptions becomes increasingly important when the missingness increases. Since we control the MAR mechanism, the assumption under which we may solve the missing data problem should hold and it is only fair to assess performance under stringent missingness conditions. We, therefore, propose to evaluate imputation methodology under 10\%, 25\% and 50\% univariate missingness (see Table \ref{table:prop}).

\begin{table}[htb]
\begin{center}
\caption{Suggested univariate missingness proportions. [TODO: change this to 'more than one missingness proportion, depending on the context' and/or add references for the given proportions (e.g. about  that 30\% is a good range for MI).]}
\label{table:prop}
\begin{tabular}{lll}
\hline
      & Reasoning for suggested missingness proportion \\
\hline  
10\%  & Depending on the size of the data, this percentage can be considered as a lower \\
      & bound of realistic evaluation. Anything less than 10\% may be of little influence on \\
      & the true data inference. Performance of a missing data method should at least be \\
      & acceptable for most missing data problems. \\ \noalign{\smallskip} % & \\
25\%  & This is a fair amount of missingness and will, depending on the observed data \\
      & information, have a noticeable influence on the completed data inference. When \\
      & compared to the condition with 10\% missingness, the inference obtained under \\
      & 25\% missingness should be less certain (i.e. confidence/credibility interval width \\
      & should increase), but estimates should still be properly covered and the statistical \\
      & properties of the missing data method should be sound. In practice, at least to our \\
      & at least to our experience in social sciences and official statistics, 25\% \\
      & univariate missingness can easily be considered as a realistic missingness percentage. \\ \noalign{\smallskip} % & \\
50\%  & Performance under 50\% percent simulated missingness will most likely be impacted \\ 
      & severely. Depending on how the missingness mechanism interacts with the simulated \\
      & data, some imputation techniques may yield estimates that are under-covered such that \\
      & the completed data inference should not be deemed valid anymore. If a method yields \\
      & acceptable inference under 50\% MAR missingness, we can determine that the statistical \\
      & properties of the imputation methodology are sound. \\
\hline
\end{tabular}
\end{center}
\end{table}

\subsection{Apply methods}

\subsection{Evaluate imputations}

@@

\subsection{Evaluate performance}

@@

\subsection{Report}

>>>>

\subsection{Study scope}

[Problem: parameters unclear, extrapolation beyond scope of simulations, misinterpretation \citep[see "investigator bias"][]{gree17}.]

\subsection{Data generation}

@@

\subsection{Missingness generation}

[TODO: change this to 'be clear which mechanisms are of interest, how they are generated, if they are suitably generated']

@@

\begin{table}[tb]
\begin{center}
\caption{Missingness mechanisms. [TODO: add refs \citet{seam13, meal15, dore18} for clarification and alternatives outlined in \citet{moha21, scho21, more18}]}
\label{table:mech}
\begin{tabular}{lll}
\hline
      & Interpretation and consequences of the mechanism \\
\hline  
MCAR  & Missing completely at random: The probability to be missing is the same for all \\
      & cases. In other words, the missing values are missing at random and the observed \\
      & values are observed at random. \\ \noalign{\smallskip}
MAR   & Missing at random: The probability to be missing is the same within groups of \\
      & cases defined by the observed data only. The essence is that observed relations in \\
      & the data inform the missingness, such that MCAR may be assumed within the \\ 
      & observed groups. \\ \noalign{\smallskip}
v-MAR & Conditional on the fully observed variables missingness occurs at random 
      & \citep[][p. 1026]{moha21}.\\
MNAR  & Missing not at random: The probability to be missing depends on unobserved \\
      & aspects of the data. The cause of the missingness is unknown and cannot be \\ 
      & inferred from the observed data. This is considered non-ignorable missingness \\
      & \citep[see e.g.][]{rubi76}. \\
\hline
\end{tabular}
\end{center}
\end{table}

@@

In addition to varying the missingness mechanisms, each simulated mechanism should be combined with different missingness patterns. Remember that missingness is only ignorable under MAR when the parameter of the data is distinct and a-priori independent from the parameter of the missing data process. Under MAR missingness we assume that we may use the observed data to make inferences about the joint (observed and unobserved) data. The dependency of the procedure on the assumption under which we obtain inference is only influenced by the amount of missingness. If there is no missingness--or if there is no data, for that matter--the inference does not depend on the assumption. Alternatively, the validity of assumptions becomes increasingly important when the missingness increases. Since we control the MAR mechanism, the assumption under which we may solve the missing data problem should hold and it is only fair to assess performance under stringent missingness conditions. We, therefore, propose to evaluate imputation methodology under several missingness proportions to amulate a realistic range in severity of the missingness problem. Depending on how the missingness mechanism interacts with the simulated data, higher missingness proportions may yield biased or invalid completed data inferences. The missingness proportions should be considered carefully.

% \begin{table}[htb]
% \begin{center}
% \caption{Suggested univariate missingness proportions. [TODO: change this to 'more than one missingness proportion, depending on the context' and/or add references for the given proportions (e.g. about  that 30\% is a good range for MI).]}
% \label{table:prop}
% \begin{tabular}{lll}
% \hline
%       & Reasoning for suggested missingness proportion \\
% \hline  
% 10\%  & Depending on the size of the data, this percentage can be considered as a lower \\
%       & bound of realistic evaluation. Anything less than 10\% may be of little influence on \\
%       & the true data inference. Performance of a missing data method should at least be \\
%       & acceptable for most missing data problems. \\ \noalign{\smallskip} % & \\
% 25\%  & This is a fair amount of missingness and will, depending on the observed data \\
%       & information, have a noticeable influence on the completed data inference. When \\
%       & compared to the condition with 10\% missingness, the inference obtained under \\
%       & 25\% missingness should be less certain (i.e. confidence/credibility interval width \\
%       & should increase), but estimates should still be properly covered and the statistical \\
%       & properties of the missing data method should be sound. In practice, at least to our \\
%       & at least to our experience in social sciences and official statistics, 25\% \\
%       & univariate missingness can easily be considered as a realistic missingness percentage. \\ \noalign{\smallskip} % & \\
% 50\%  & Performance under 50\% percent simulated missingness will most likely be impacted \\ 
%       & severely. Depending on how the missingness mechanism interacts with the simulated \\
%       & data, some imputation techniques may yield estimates that are under-covered such that \\
%       & the completed data inference should not be deemed valid anymore. If a method yields \\
%       & acceptable inference under 50\% MAR missingness, we can determine that the statistical \\
%       & properties of the imputation methodology are sound. \\
% \hline
% \end{tabular}
% \end{center}
% \end{table}

\subsection{Imputation methods}

\subsection{Evaluation}

@@

@@

@@

@@

\subsection{Reporting}
